Sparse Function-space Representation \\ of Neural Networks

### 2023-05-14 paper/main.tex
rewrite 2.1 first draft done

Although the true posterior is intractable given the non-linearities of the neural network, one can restore to sampling techniques such as Hamiltonian Monte-Carlo (HMC). However, even in this case, the high computational costs make them impractical. In \cref{fig:banana}, we show a qualitative example of the induced function space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset. \todo{make it more straightforward that we are computationally cheaper?}

A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function  
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$ 
with the Generalized Gaussian-Newton approximation (GGN). 
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}) = \Jac{\weights_*}{\mathbf{x}} \, \weights$, where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian at the MAP. Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$. As pointed out in \citet{khan2019approximate}, a Gaussian posterior covariance approximation in the weight space implies a covariance structure in the function space. \todo{the last two sentences are not linked to the previous ones}

In this paper, we are concerned with learning the posterior $q(\vf)$ directly in the function space, instead of $q(\vw)$. We choose to use a Gaussian process, an extension of the multivariate normal that characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance  (or kernel) function $\kappa(\cdot,\cdot)$. We aim to these two functions based on the neural network to then build the Gaussian process posterior. A function linear with respect to its parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows: \todo{this part needs to be improved the linear function pops out of nowhere}
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = 0 \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}

Therefore, we can use the approximate linear model defined by the Laplace-GGN approximation to define our GP derived from the trained neural network,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\JacT{\weights^*}{\mathbf{x}'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual data, they adjusted the data to be pseudo-data.\todo{The reader may wonder what is pseudo-data then} Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the functional representation of our trained neural network. \todo{this sentence can be avoided probably}

>>>>

Although the true posterior is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC). However, even in this case, the high computational costs make them impractical meaning for most applications we need approximate inference techniques. In \cref{fig:banana}, we show a qualitative example of the induced function space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset. \todo{make it more straightforward that we are computationally cheaper?}

A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function  
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$ 
with the Generalized Gaussian-Newton approximation (GGN). 
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}) = \Jac{\weights_*}{\mathbf{x}} \, \weights$, where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian at the MAP. Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.

In this paper, we are concerned with learning the posterior $q(\vf)$ directly in the function space, instead of $q(\vw)$. We choose to use a Gaussian process, an extension of the multivariate normal that characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance  (or kernel) function $\kappa(\cdot,\cdot)$. We aim to derive these two functions based on the neural network to then build the Gaussian process posterior. Gaussian distributions remain tractable under learn transformations. Therefore, a linear function with respect to its parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows: \todo{this part needs to be improved the linear function pops out of nowhere}
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = 0 \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}

Therefore, the GGN approximate linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\JacT{\weights^*}{\mathbf{x}'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual data, they adjusted the data to be pseudo-data.\todo{The reader may wonder what is pseudo-data then} Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the functional representation of our trained neural network. \todo{this sentence can be avoided probably}
